---
title: "Statistical Inference Course Project"
author: "Edilmo Palencia"
date: "August 22, 2015"
output: 
  html_document: 
    theme: cerulean
---

# Tooth Growth Analysis

The data present the next structure:  
```{r, echo=FALSE}
str(ToothGrowth)
```
  
We have 60 observations with tree variables: length is numeric, supplement is a factor and dose is numeric as well.  
  
A quick summary of the data:  
```{r, echo=FALSE, results='asis'}
knitr::kable(summary(ToothGrowth))
```
  
In the table we can see:  
- There are 30 observation for each dosis.  
- The minimum, maximun, mean, median, first and third quantile for the length and the doses.  
  
Let's see how many observations we have per sup and types of doses:  
```{r echo=FALSE}
t<-ToothGrowth
t[3]<-sapply(t[[3]],as.character)
t[3]<-factor(t[[3]])
ftable(t[,2:3])
```
  
Let's see the distributions of the length per sup and dose:  
```{r echo=FALSE}
library(ggplot2)
g.dis <- ggplot(data = t, aes(x = len))
g.dis <- g.dis + geom_histogram(alpha = .20, binwidth=.3, colour = "black", aes(y = ..density..))
g.dis <- g.dis + facet_grid(supp ~ dose) + ggtitle("Distribution of len for sup & dose")
g.dis
```

Let's compute the t-test over len for each group of sup and dose:  
```{r echo=FALSE}
t.groups <- split(t[[1]],list(t[[2]],t[[3]]))
t.groups.test <- sapply(t.groups,function(x){t.test(x)[c("conf.int","estimate","p.value")]})
t.groups.test <- t.groups.test[c(1,1,2,3),]
t.groups.test[c(1,2),]<-sapply(t.groups.test[1,],as.numeric)
t.groups.test
```
  
In the above table we can se for each group of sup and dose: the sample mean, the t confidence interval those means, and the p values.

## Asumptions

From the data presented above we precibe that the sup of value OJ is more efective than VC. To support this hipotesis let's compare the means.

```{r echo=FALSE}
OJ.means<-as.vector(t.groups.test[3,c(1,3,5)],mode = "numeric")
VC.means<-as.vector(t.groups.test[3,c(2,4,6)],mode = "numeric")
t.test(OJ.means,VC.means, var.equal = FALSE, paired = TRUE)
```
  
In the results presented we can see that the 95% confidence interval includes 0, with a p-value of .19 that is greater than .05  

## Conclusions

There are few samples to confirm the assumption stated above. This conclussion is derived from the fact that the confidence interval includes the 0 and it is quite wide.
